---
title: "Exploring Governance Model Distributions"
author: "Eric Stone"
date: '2022-08-31'
output: html_document
---

## Demonstrating R + ShroomDK

We are going to use histograms, density plots, and a modified 
Receiver Operating Characteristic (or ROC) curve to understand and compare 
the wealth distributions of a few popular ERC-20 and NFT-based tokens.

The idea here is to understand whether one model or the other results in 
better setups for more egalitarian governance.

This part is a bit arbitrary, but as you'll see, by using [@flipsidecrypto](https://twitter.com/flipsidecrypto)'s
ShroomDK API, it's incredibly easy for you to fork this, add other projects
and get more creative with the visualizations and stats. 

For example: This R document is available to fork in flipside's public tools repo. Have at it 🤓s!
[github.com/FlipsideCrypto/tools/tree/main/public-examples](https://github.com/FlipsideCrypto/tools/tree/main/public-examples)

I'll walk you thru how to set it up in R below, but you can go to [sdk.flipsidecrypto.xyz](https://sdk.flipsidecrypto.xyz/shroomdk)
to mint your API key and get started first if you want to follow along.

For this exercise I'm going to use 3 active NFT projects with different floors (
moonbirds, nouns, and bgans), and compare to 3 prominent governance tokens for
major defi projects (1inch, aave, and comp).

FIRST SOME SETUP

```{r setup, include=FALSE, echo=TRUE}
### ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ ####
### THIS requires a free flipside api key to run queries ###
### ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ ####

# flipside crypto's massive open dataset via R
# (free to all: https://sdk.flipsidecrypto.xyz/shroomdk)
# follow the setup steps for the R SDK 
#   -> here https://docs.flipsidecrypto.com/shroomdk-sdk/sdks/r
# once you have that set up this should run!

# 🐍 there's a package for python too.. we bilingual
# https://docs.flipsidecrypto.com/shroomdk-sdk/sdks/python

#load required packages
#one-time setup for shroomDK
if(FALSE){
  library(devtools) # install if you haven't already
  devtools::install_github(repo = 'FlipsideCrypto/sdk', subdir = 'r/shroomDK')
  #remember to save your api key to GIT IGNORED location
  #and create an object to refer to it
  api_key <- readLines("shroomdk.key")
  
  #this helper function is the most straightforward way to query the database
  #  --> shroomDK::auto_paginate_query(query, api_key) # <- pulls up to 1M rows of a query & attempts to make it a nice data frame. 
  #explore the data in a sql editor at https://flipside.new
}

library(shroomDK)
library(magrittr)
library(sde)
library(sn)
library(grid)
library(quantmod)
library(data.table)
library(ggplot2)
api_key <- readLines("shroomdk.key") 
options(scipen = 999)

color.pallet <- c("#1AB9FF", "#D17DFE", "#53E1AE",
                  "#FCC003", "orange", "#DCF180", 
                  "#63528C", "pink")
```


### Get the data and play around a bit

Now let's get some data into our environment and see what we can see!

```{r collect_explore}
#our own table to keep track of the projects we're observing/comparing
contracts <- data.table(
  project_name = c("moonbirds","nouns","bgans",
                   "aave","comp","1inch"),
  type = c("nft","nft","nft","gov","gov","gov"),
  contract_address = c(
    tolower('0x23581767a106ae21c074b2276D25e5C3e136a68b'),
    tolower('0x9C8fF314C9Bc7F6e59A9d9225Fb22946427eDC03'),
    tolower('0x31385d3520bCED94f77AaE104b406994D8F2168C'),
    tolower('0x7Fc66500c84A76Ad7e9c93437bFc5Ac33E2DDaE9'),
    tolower('0xc00e94Cb662C3520282E6f5717214004A7f26888'),
    tolower('0x111111111117dC0aa78b770fA6A738034120C302')
  )
)

#let's consider any holder that has been active in the last 6 months
holder.data <- data.table(
  shroomDK::auto_paginate_query(
    paste0(
      "select *
      from ethereum.core.ez_current_balances
      where last_activity_block_timestamp > current_date - (6 * 31)
      and contract_address in ('",paste0(
        contracts$contract_address, collapse = "','"
      ),"');"
    ),
    api_key
  )
)

#here's a similar query in the flipside app (obviously not able to reference our contracts table)
# https://app.flipsidecrypto.com/velocity/queries/d0720e7e-df2b-430b-a82b-62c6b78c1cd1

#short cleanup process
setnames(holder.data, tolower(names(holder.data))) #because i don't like uppercase names
holder.data[ contract_address %in% contracts[ type == "nft" ]$contract_address,
             current_bal := current_bal_unadj ] #for NFTs the unadjusted balance is the number of holders
holder.data <- merge(
  holder.data,
  contracts,
  by = "contract_address"
) #make sure i've got the right names and classes going

# holy cow now we can look at histograms what is this
ggplot(holder.data, aes(x = current_bal, fill = type)) + 
  geom_histogram( bins = 20 ) +
  facet_wrap( ~ project_name, scales = "free_x", ncol = 2 ) +
  scale_fill_manual(values = color.pallet) +
  theme_light() +
  theme(legend.position = "top")

#this makes me want to look at governance vs nft first
ggplot(holder.data, aes(x = current_bal, fill = type)) + 
  geom_histogram( bins = 20 ) +
  facet_wrap( ~ type, scales = "free_x", ncol = 2 ) +
  scale_fill_manual(values = color.pallet) +
  theme_light() +
  theme(legend.position = "top")

#as expected most holders hold small amounts with long tails
# of larger accounts

#this is more pronounced with our governance tokens than NFTs

#but to get a sense of the shapes we need to look at 
#a log-adjusted x axis

ggplot(holder.data[ type == 'gov' ],aes(x = current_bal)) + 
  geom_histogram( bins = 20 ) +
  facet_wrap( ~ project_name, scales = "fixed", ncol = 1 ) +
  theme_light() +
  scale_x_log10() +
  scale_fill_manual(values = color.pallet) +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
  )

#now that we zoom in, we clearly need to eliminate tiny amounts
#even a floor of .01 tokens gives us a better view

ggplot(holder.data[ type == 'gov' & current_bal > .01 ],
       aes(x = current_bal, fill = project_name)) + 
  geom_density( alpha = .4 ) +
  theme_light() +
  scale_x_log10() +
  scale_fill_manual(values = color.pallet) +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
  )

ggplot(holder.data[ type == 'nft' & current_bal > .01 ],
       aes(x = current_bal, fill = project_name)) + 
  geom_density( alpha = .4 ) +
  theme_light() +
  scale_x_log10() +
  scale_fill_manual(values = color.pallet) +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
  )


#this is more helpful
#comp clearly has fewer active 6 month holders than aave
#confirm:
holder.data[ type == "gov" & current_bal > .01,
             list(six_mo_active_holders = .N),
             by = project_name ]

#but their distribution shapes are uncannily similar! not a huge surprise

#what is perhaps more interesting is the holder distributions of 
#NFTs-focused-on-community-participation vs. two active governance tokens
#for significant lending protocols

#now we'll see some differences..
ggplot(holder.data[ current_bal > .01 & current_bal < 100 ],
       aes(x = current_bal, fill = type)) + 
  geom_density( alpha = .4 ) +
  theme_light() +
  scale_x_log10() +
  scale_fill_manual(values = color.pallet) +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
  )
```

## Going Deeper

Now that we've had a look around we realize we need a couple things to do this
properly, namely:

1. Percentage based comparisons of supply distribution (absolutes are all over the place here)
2. Address Labels (to account for non-user-controled tokens as best as possible)

```{r deepR}

#what do we actually care about when comparing these two?
#right! percent of the supply owned by active addresses
holder.data[
  ,percent_supply := current_bal / sum(current_bal),
  by = contract_address ]

#ideally, you'd want a) more individual participants
#and b) a "shorter-tailed" holder distribution
#for any of these projects
#and from this view it would appear the NFT projects have a significant edge!

#but given the annoying shapes of these distributions, let's plot them a different way
#cumulative percent of supply against cumulative percent of users:
#the closer this curve to a 45 degree line the better
#in terms of egalitarian token distribution

#so let's get cumulative ownership percentages calculated:
setorderv(holder.data, cols = c("contract_address", "percent_supply"), order = c(1L,-1L))
holder.data[,supply_cumulative := cumsum(percent_supply), by = contract_address ]
holder.data[,users_cumulative := 1:.N / .N, by = contract_address ]

ggplot(
  holder.data,
  aes(y = supply_cumulative, x = users_cumulative,
           color = project_name, 
           linetype = type)) +
  scale_color_manual(values = color.pallet) +
  geom_line( size = 1 ) +
  theme_light()

#what else do we need? address labels so that we're eliminate treasury
#and exchange hot/cold wallets, pools, and reserves from both sets

#let's subset to addresses that hold meaningful amounts of the tokens
addy.labels <- data.table(
  shroomDK::auto_paginate_query(
    paste0(
      "select *
      from ethereum.core.dim_labels
      where address in ('",paste0(
        unique(c(
          holder.data[ type == 'nft' & current_bal > 3 ]$user_address,
          holder.data[ type == 'gov' & percent_supply > .001 ]$user_address
        )), collapse = "','"
      ),"');"
    ),
    api_key
  )
)
setnames(addy.labels, tolower(names(addy.labels)))

holder.data <- merge(
  holder.data,
  addy.labels[ ,list(address,address_name,label_subtype)],
  by.x = "user_address",
  by.y = "address",
  all.x = TRUE
)

holder.data.adj <- holder.data[(address_name == 'gnosis safe: general contract' |
                                  is.na(address_name))][ user_address != "0x26a78d5b6d7a7aceedd1e6ee3229b372a624d8b7" ] #clearly a kraken address
setorderv(holder.data.adj, cols = c("contract_address", "percent_supply"), order = c(1L,-1L))
holder.data.adj[ ,percent_supply := current_bal / sum(current_bal), by = contract_address ]
holder.data.adj[ ,supply_cumulative := cumsum(percent_supply), by = contract_address ]
holder.data.adj[ ,users_cumulative := 1:.N / .N, by = contract_address ]

ggplot(
  holder.data.adj,
  aes(y = supply_cumulative, x = users_cumulative,
      color = project_name, 
      linetype = type)) +
  scale_color_manual(values = color.pallet) +
  geom_line( size = 1 ) + 
  ggtitle("Distribution of Supply as Percentages (labeled addys removed)") +
  theme_light()

```


### Wrapping up for now

There's obviously a lot more we can do here, but part of the point is I'm
just scratching the surface, and putting this somehwere you can fork it,
recreate it with your own ShroomDK key (free!), and do what you like. 

Expand, build, ignore, whatever.

In the last snippet here I'm going to calculate the respective areas under
each curve and look at a couple different metrics to summarize holders.

IF YOU WANTED TO GO DEEPER
the next logical step is to incorporate actual governance participation and voting
(worth noting that flipside has Snapshot data in its DB as well)

```{r summary}
library(zoo)

contracts$auc_supply <- unlist(lapply(contracts$project_name, function(x){
  sum(
    diff( holder.data.adj[ project_name == x ]$users_cumulative ) *
      rollmean( holder.data.adj[ project_name == x ]$supply_cumulative, 2 )
  )
}))

contracts$n_control_95 <- unlist(lapply(contracts$project_name, function(x){
  nrow(holder.data.adj[ project_name == x & supply_cumulative <= .95 ])
})) 
contracts
```
Perhaps the most unexpected takeaway I have from this EDA is how good it makes moonbirds look.
Kudos to them, but that floor.. is too.. high... innit.

[@theericstone](https://twitter.com/theericstone) if you like

